---
title: "ML API Specification for EO Data Cubes"
author: "Brian Pondi and Rolf Simoes"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Temporal CNN Example

This section demonstrates how to train a Temporal CNN (TempCNN) model using OpenEO processes. The example uses deforestation data from Rondonia to train a deep learning model for time series classification.

## Setup and Connection

First, we load the required libraries and connect to the OpenEO backend:

```{r}
library(openeo)

# Connect to the backend
con <- connect(
    host = "http://127.0.0.1:8000",
    user = "brian",
    password = "123456"
)

# Access processes
p <- processes()
```
## Describe processes
```{r}
openeo::describe_process("mlm_class_tempcnn")
```


```{r}
openeo::describe_process("mlm_class_random_forest")
```

## Load Training Data

We load the pre-processed training data for deforestation in Rondonia:

```{r}
# Load sits tibble Training data
data_deforestation_rondonia <- readRDS("./data/monthly_rondonia_data.rds")
```

## Define the Temporal CNN Model

Here we define the architecture and hyperparameters for our TempCNN model:

```{r}
# Define the Temporal CNN model
tempcnn_model_init <- p$mlm_class_tempcnn(
    optimizer = "adam",
    learning_rate = 0.0005,
    seed = 42
)
```

## Model Training

Now we fit the model using our training dataset:

```{r}
# Fit the model using the training dataset
tempcnn_model <- p$ml_fit(
    model = tempcnn_model_init,
    training_set = jsonlite::serializeJSON(data_deforestation_rondonia),
    target = "label"
)
```

## Save Model

```{r}
# Save the model
model <- p$save_ml_model(
    data = tempcnn_model,
    name = 'tempcnn_rondonia_model',
    tasks = list("classification")
)
```

## Load Sentinel-2 Data

```{r}
datacube <- p$load_collection(
    id = "mpc-sentinel-2-l2a",
    spatial_extent = list(
        west = -63.9,
        east = -62.9,
        south = -9.14,
        north = -8.14
    ),
    bands = list(
        "B02",
        "B03",
        "B04",
        "B05",
        "B06",
        "B07",
        "B08",
        "B11",
        "B12",
        "B8A"
    ),
    temporal_extent = c("2022-01-01", "2022-12-31")
)
```

# Regularize the data

```{r}
datacube <- p$cube_regularize(
    data = datacube,
    period = "P1M",
    resolution = "320"
)
```

# Apply the trained model to make predictions


```{r}
data <- p$ml_predict(
    data = datacube,
    model = tempcnn_model
)
```

## # Save the prediction results

```{r}
ml_job <- p$save_result(
    data = data,
    format = "GTiff"
)
```
# Run the job
```{r}
job <- create_job(
    graph = ml_job,
    title = "Temp CNN Model - Rondonia - R Client",
    description = "Training a Temp CNN model, Predict with it"
)
job <- start_job(job)

# Display job information
describe_job(job)
```

## Conclusion

This example demonstrated how to:
1. Connect to an OpenEO backend
2. Load and prepare training data
3. Define a Temporal CNN model architecture
4. Train the model
5. Make predictions on new data
6. Save the results
